What We Eval in the Shadows

### 2021-08-12 paper/main.tex
Proofread of 2.3

The expression passed to \eval can be thought of as an abstract syntax
tree. It can be parsed from a string, created manually or using reflection.

\begin{figure}[htb]
\begin{minipage}{.49\textwidth}
\begin{lstlisting}[caption={Examples of calls producing expression \k{a+b}}]
  parse(text="a+b")
  quote(a+b)
  call("+", quote(a), quote(b))

  # reflect promise
  f <- function(x) substitute(x)
  f(a+b)
  \end{lstlisting}
\end{minipage}
\begin{minipage}{.49\textwidth}
  \begin{lstlisting}[caption={Example of a call reflection}]
  f <- function(x, y) {
    mc <- match.call() # reflect curr. call
    mc[[1]] <- as.name("g")
    mc[["x"]] <- 2
    mc
  }
  f(1,2) # returns a call g(2,y)
  \end{lstlisting}
\end{minipage}
\end{figure}

 The call to \k{substitute(x)} extracts the unevaluated expression from
the promise \k{x}. The function \k{match.call} returns an expression
representing the current call.

R is generous when it comes to what is allowed to be passed as environment to
\eval. Besides ordinary environments, it also accepts a list or a data
frame (using element names or column names for variable resolution) or an
integer \k{n} (in which case it would use the \k{n.} call frame). If no
environment is specified, then the default is the environment where the call to
\eval was made (as returned by \k{parent.frame}).
%
Environments nest, each has a parent. When creating a new environment with
\k{new.env}, the parent is the current environment. Environment chains can be
traversed with \k{parent.env}, until \k{emptyenv} is reached. The top-level
environment is \k{.GlobalEnv}, it has parents that represent the packages that
have been loaded. One can also directly read, modify or create new bindings,
given any environment:
%
\begin{lstlisting}
  # reading                                # writing
  envir$v                                  envir$v <- 2
  get("v", envir=envir)                    assign("v", 2, envir=envir)
\end{lstlisting}
%
 Environments are used as hash maps as they have reference semantics
and a built-in string lookup.

>>>>

The expression passed to \eval can be thought of as an abstract syntax tree.
Listings~\ref{lst:exprs} and \ref{lst:match.call} show some of the ways of
creating expressions: parsing from a string, manually or via reflection. The
call to \k{substitute(x)} extracts the unevaluated expression from the promise
\k{x}. The call to \k{match.call()} returns an expression representing the
current call which can then be further modified.

\begin{figure}[htb]
\begin{minipage}{.49\textwidth}
\begin{lstlisting}[caption={Examples of calls producing expression \k{a+b}},label=lst:exprs]
  parse(text="a+b")
  quote(a+b)
  call("+", quote(a), quote(b))

  # reflect promise
  f <- function(x) substitute(x)
  f(a+b)
  \end{lstlisting}
\end{minipage}
\begin{minipage}{.49\textwidth}
  \begin{lstlisting}[caption={Example of a call reflection},label=lst:match.call]
  f <- function(x, y) {
    mc <- match.call() # reflect curr. call
    mc[[1]] <- as.name("g")
    mc[["x"]] <- 3
    mc
  }
  f(1,2) # returns an expr g(x=3,y=2)
  \end{lstlisting}
\end{minipage}
\end{figure}

R is generous when it comes to what is allowed to be passed as environment to
\eval. Besides ordinary environments, it also accepts a list or a data
frame (using element names or column names for variable resolution) or an
integer \k{n} (in which case it would use the \k{n.} call frame). If no
environment is specified, then the default is the environment where the call to
\eval was made (as returned by \k{parent.frame}).
%
Environments nest, each has a parent. When creating a new environment with
\k{new.env}, the parent is the current environment. Environment chains can be
traversed with \k{parent.env}, until \k{emptyenv} is reached. The top-level
environment is \k{.GlobalEnv}, it has parents that represent the packages that
have been loaded. One can also directly read, modify or create new bindings,
given any environment:
%
\begin{figure}[H]
\begin{minipage}{.49\textwidth}
  \begin{lstlisting}
  # reading
  envir$v              
  get("v", envir=envir)
  \end{lstlisting}
\end{minipage}
\begin{minipage}{.49\textwidth}
  \begin{lstlisting}
  # writing
  envir$v <- 2
  assign("v", 2, envir=envir)  
  \end{lstlisting}
\end{minipage}
\end{figure}
%
\noindent Environments are used as hash maps as they have reference semantics
and a built-in string lookup.
